Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2015-11-15 manuscript.tex
Add 2 other refs for shell effect
Talk about analytical solutions in intro
Before saying that need to solve numerically, say that there are
analytical in special cases of a spherical cap

Satellite missions dedicated to measuring the Earth's gravity field
(like CHAMP, GRACE, and GOCE)
have provided geophysicists with almost uniform and global data coverage.
These new data have enabled interpretations on regional and global scales
\citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
Modeling at such scales requires taking into account the curvature of the
Earth and calculating gravity gradients as well as the traditional
gravitational acceleration.
A common approach to achieve this is
to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
instead of rectangular prisms.
However, the integral formula for the gravitational effects of a tesseroid
must be solved numerically.
Approaches to this numerical integration include
Taylor series expansion \citep{Heck2007, Grombein2013}
and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
Taylor series expansion produces accurate results at low latitudes but
presents a decrease in accuracy towards the polar regions.
This is attributed to tesseroids degenerating into an approximately triangular
shape at the poles.
The Gauss-Legendre Quadrature (GLQ) integration consists in approximating
the volume integral by a weighted sum of the effect of point masses.
An advantage of the GLQ approach is that it can be
controlled by the number of point masses used.
The larger the number of point masses,
the better the accuracy of GLQ integration.
A disadvantage is the increased computation time as the number of
point masses increases.
Thus, there is a trade-off between accuracy and computation time.
This is a common theme in numerical methods.
\citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
including tesseroids, to compute the gravitational effects of topographic
masses.
The author concludes that using tesseroids with GLQ integration gives the best
results for near-zone computations.
However, the question of how to determine the optimal parameters for GLQ
integration remained open.

@@

The reference against which we compare the computed tesseroid fields
is a homogeneous spherical shell.
The shell has analytical solutions along the polar axis
\citep{Grombein2013}
and can be perfectly discretized into tesseroids.
We chose a spherical shell with a thickness of 1 km,
density of $2670\ kg.m^{-3}$,
bottom at height 0 km above the reference sphere,
and top at 1 km height.
We produced tesseroid models of the shell by discretizing it along the
horizontal dimensions into a regular mesh.

>>>>

Satellite missions dedicated to measuring the Earth's gravity field
(like CHAMP, GRACE, and GOCE)
have provided geophysicists with almost uniform and global data coverage.
These new data have enabled interpretations on regional and global scales
\citep[e.g.][]{Reguzzoni2013,Braitenberg2015}.
Modeling at such scales requires taking into account the curvature of the
Earth and calculating gravity gradients as well as the traditional
gravitational acceleration.
A common approach to achieve this is
to discretize the Earth into tesseroids (Figure~\ref{fig:tesseroid})
instead of rectangular prisms.
An analytical solution exists when the computation point is along
the polar axis and the tesseroid is extended into a spherical cap
\citep{Lafehr1991, Mikuska2006, Grombein2013}.
For more general cases,
the integral formula for the gravitational effects of a tesseroid
must be solved numerically.
Approaches to this numerical integration include
Taylor series expansion \citep{Heck2007, Grombein2013}
and the Gauss-Legendre Quadrature \citep{Asgharzadeh2007}.
Taylor series expansion produces accurate results at low latitudes but
presents a decrease in accuracy towards the polar regions.
This is attributed to tesseroids degenerating into an approximately triangular
shape at the poles.
The Gauss-Legendre Quadrature (GLQ) integration consists in approximating
the volume integral by a weighted sum of the effect of point masses.
An advantage of the GLQ approach is that it can be
controlled by the number of point masses used.
The larger the number of point masses,
the better the accuracy of GLQ integration.
A disadvantage is the increased computation time as the number of
point masses increases.
Thus, there is a trade-off between accuracy and computation time.
This is a common theme in numerical methods.
\citet{Wild-Pfeiffer2008} investigated the use of different mass elements,
including tesseroids, to compute the gravitational effects of topographic
masses.
The author concludes that using tesseroids with GLQ integration gives the best
results for near-zone computations.
However, the question of how to determine the optimal parameters for GLQ
integration remained open.

@@

The reference against which we compare the computed tesseroid fields
is a homogeneous spherical shell.
The shell has analytical solutions along the polar axis
\citep{Lafehr1991, Mikuska2006, Grombein2013}
and can be perfectly discretized into tesseroids.
We chose a spherical shell with a thickness of 1 km,
density of $2670\ kg.m^{-3}$,
bottom at height 0 km above the reference sphere,
and top at 1 km height.
We produced tesseroid models of the shell by discretizing it along the
horizontal dimensions into a regular mesh.
